---
title: "Preprocess the IDS data"
output: html_notebook
---

# Prepare data

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

datadir <- "../data/cicids17.csv"
ids_data <- read_csv(datadir)
```

## Check for and handle null/inf values

```{r}
# get count of nulls
ids_null <- map(ids_data, ~sum(is.na(.)))

# identify cols with null
col_with_null <- which(ids_null > 0)
print("Cols with null")
print(ids_null[col_with_null])
```

We find that `flow bytes` is the only column with missing values
Let's take a closer look at the column

```{r}
null_col <- ids_data[col_with_null]
summary(null_col)
```

Let's try to find what type of data points has null values

```{r}
null_row <- which(is.na(null_col))
null_df <- ids_data %>% slice(null_row)

qplot(Label, data = null_df, main = "Distribution of rows with null values across labels")

ids_data %>% select(Label) %>% filter(Label == "DoS Hulk") %>% count(Label) %>% print
```

Since the no of observations with nulls are significantly lower we drop the rows with null values

```{r}
ids_data <- ids_data %>% drop_na()
write_csv(ids_data, "../data/cicids17_clean.csv")  # save current state
```

## Handling inf values

```{r}
datadir <- "../data/cicids17_clean.csv"
ids_data <- read_csv(datadir)
```

Get infinity count

```{r}
# get count of inf
ids_inf <- map(ids_data, ~sum(is.infinite(.)))

# identify cols with inf
col_with_inf <- which(ids_inf > 0)
print("Cols with inf")
print(ids_inf[col_with_inf])
```

```{r}
inf_col <- ids_data[col_with_inf]
summary(inf_col)
```

Let's try to find what type of data points has inf values

```{r}
inf_row <- which(is.infinite(inf_col[[2]]))
inf_df <- ids_data %>% slice(inf_row)

qplot(Label, data = inf_df, main = "Distribution of rows with inf values across labels")
```

Dropping all the rows with infinite values

```{r}
clean_df <- ids_data %>% filter(is.finite(`Flow Bytes/s`) & is.finite(`Flow Packets/s`))
write_csv(clean_df, "../data/cicids17_clean.csv")
```
